---
title: "Cpts 411 Programming Project 4"
author: 
     - "Clancy Andrews"
     - "Alex Shirk"
header-includes:
   - \usepackage{amssymb}
   - \usepackage{amsmath}
   - \usepackage{graphicx,float}
   - \usepackage{natbib}
   - \usepackage{geometry}
   - \usepackage{xcolor}
   - \usepackage{courier}
output:
  pdf_document: default
fontsize: 11pt
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\break

# Introduction

|       For this project, our goal was to use a multithreaded program to approximate the value of $\pi$. Given a unit square with center at $(0.5,0.5)$, overlay a unit circle in the unit square with origin at $(0.5,0.5)$. From there, we can randomly select a coordinate inside the unit square (the same idea as throwing a dart at the square). If the coordinates are located inside the unit circle, then we can increment the number of hits by one. If the coordinates are not located inside the unit circle, then we do not increment the number of hits. It is worth noting that the coordinates are always located inside the unit square. Letting $n$ be the number of random coordinates generated and $h$ be the number of those generated coordinates also being inside the unit circle, we can use the following equation to approximate $\pi$:
\begin{align}
\pi = \frac{4h}{n}
\end{align}
The data was collected on a personal computer, which had an Intel i9-10900K CPU at 3.70GHz, containing 10 cores and 20 logical processors (threads).

\bigbreak

# Analysis

|       For comparing the estimated value of $\pi$ with the actual value, we used $$\pi = 3.141592653589793238462643383279502884197$$ as the actual value. We got this number from the following webpage: [https://www.britannica.com/science/pi-mathematics](https://www.britannica.com/science/pi-mathematics). Below we can observe the speedup table where the values of $n$ are the row labels and the values of $p$ are the column labels.

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Libraries
library(ggplot2)
library(knitr)
library(stats)

#Import data for analysis
df = read.csv2("output.csv", header = FALSE, sep = ",")
colnames(df) = c("Pi","P", "N", "Difference", "Time")

#Extract Serial Time from Data
serial_time = subset(df, P == 1)$Time

#Get the parallel times for each P value
parallel_time = data.frame(matrix(ncol = length(unique(df$P)) - 1, nrow = 11))
colnames(parallel_time) = c("1", "2", "4", "8")

for (i in colnames(parallel_time)) {
  p_value = as.numeric(i)
  data = df$Time[df$P == p_value]
  parallel_time[, i] = data
}

#Calculate the speed up
speedup= data.frame(matrix(ncol = length(unique(df$P)) - 1, nrow = 11))
colnames(speedup) = c("1", "2", "4", "8")

for (i in colnames(speedup)) {
  speedup[,i] = as.numeric(serial_time)/as.numeric(parallel_time[,i])
}

row.names(speedup) = unique(df$N)

#Precision data
pi = data.frame(matrix(ncol = length(unique(df$P)), nrow = 11))
colnames(pi) = c("1", "2", "4", "8", "16")

for (i in colnames(pi)) {
  p_value = as.numeric(i)
  data = df$Pi[df$P == p_value]
  data = round(as.numeric(data),5)
  pi[, i] = data
}
row.names(pi) = unique(df$N)

#Difference in the pi values
pi_diff = data.frame(matrix(ncol = length(unique(df$P)), nrow = 11))
colnames(pi_diff) = c("1", "2", "4", "8", "16")

for (i in colnames(pi)) {
  p_value = as.numeric(i)
  data = df$Difference[df$P == p_value]
  data = round(as.numeric(data),5)
  pi_diff[, i] = data
}
row.names(pi_diff) = unique(df$N)
```



```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(speedup, row.names = TRUE, caption = "Speed up for corrsponding $n$ (rows) and $p$ (columns) values.")
```

Using our previously presented true value of $\pi$ to compare, we can observe the estimated values of $\pi$ approximated by our program in the table below.

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(pi, row.names = TRUE, caption = "Estimations for the value of $\\pi$ for corrsponding $n$ (rows) and $p$ (columns) values.")
```
From observation, we can see that as $n$ increased in size, our approximations would get closer to the true value of $\pi$. The next table shows the difference between the true and estimated $\pi$ values at each corresponding $n$ and $p$ value.

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(pi_diff, row.names = TRUE, caption = "Differences in the true value versus the estimated value of $\\pi$ for corrsponding $n$ (rows) and $p$ (columns) values.")
```
It is more evident now that as we increase $n$ and $p$ together, we get a closer approximation to the true value of $\pi$.

\break
# Analysis Code

The following is the code used for analyzing the collected data from the program:

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
#Libraries
library(ggplot2)
library(knitr)
library(stats)

#Import data for analysis
df = read.csv2("output.csv", header = FALSE, sep = ",")
colnames(df) = c("Pi","P", "N", "Difference", "Time")

#Extract Serial Time from Data
serial_time = subset(df, P == 1)$Time

#Get the parallel times for each P value
parallel_time = data.frame(matrix(ncol = length(unique(df$P)) - 1, nrow = 11))
colnames(parallel_time) = c("1", "2", "4", "8")

for (i in colnames(parallel_time)) {
  p_value = as.numeric(i)
  data = df$Time[df$P == p_value]
  parallel_time[, i] = data
}

#Calculate the speed up
speedup= data.frame(matrix(ncol = length(unique(df$P)) - 1, nrow = 11))
colnames(speedup) = c("1", "2", "4", "8")

for (i in colnames(speedup)) {
  speedup[,i] = as.numeric(serial_time)/as.numeric(parallel_time[,i])
}

row.names(speedup) = unique(df$N)

#Precision data
pi = data.frame(matrix(ncol = length(unique(df$P)), nrow = 11))
colnames(pi) = c("1", "2", "4", "8", "16")

for (i in colnames(pi)) {
  p_value = as.numeric(i)
  data = df$Pi[df$P == p_value]
  data = round(as.numeric(data),5)
  pi[, i] = data
}
row.names(pi) = unique(df$N)

#Difference in the pi values
pi_diff = data.frame(matrix(ncol = length(unique(df$P)), nrow = 11))
colnames(pi_diff) = c("1", "2", "4", "8", "16")

for (i in colnames(pi)) {
  p_value = as.numeric(i)
  data = df$Difference[df$P == p_value]
  data = round(as.numeric(data),5)
  pi_diff[, i] = data
}
row.names(pi_diff) = unique(df$N)

#Tables for speedup and pi_estimates
kable(speedup, row.names = TRUE, caption = "Speed up for corrsponding $n$ (rows) and $p$ (columns) values.")

kable(pi, row.names = TRUE, caption = "Estimations for the value of $\\pi$ for corrsponding $n$ (rows) and $p$ (columns) values.")

kable(pi_diff, row.names = TRUE, caption = "Differences in the true value versus the estimated value of $\\pi$ for corrsponding $n$ (rows) and $p$ (columns) values.")
```


\break
# Session Info
```{r}
sessionInfo()
```

